---
title: "red-pandas"
author: "Zak Varty"
format: html
---

# Red Pandas 

The red panda (*Ailurus fulgens*) is a small, tree-dwelling mammal native to the Himalayan region and parts of south-western China. Despite sharing a name with the giant panda, the red panda is not closely related; it belongs to its own unique family, Ailuridae. With its russet-coloured fur, bushy ringed tail, and expressive face, the red panda is often considered one of the most charismatic animals in the world. 

## Appearance

Red pandas are about the size of a domestic cat, with a body length of 50–65 cm and a tail adding another 30–50 cm. They typically weigh between 3–6 kilograms. 

Their thick reddish-brown fur provides insulation against cold mountain climates, while the distinctive white markings on their face help with individual recognition. 

## Habitat and Diet 

These animals inhabit temperate forests at elevations of 2,200–4,800 meters, where dense bamboo understories are common. Bamboo makes up the majority of their diet, although they also eat fruits, berries, acorns, and occasionally insects or small vertebrates. Because bamboo is low in nutrients, red pandas spend much of their time feeding and resting. 

## Behavior 

Red pandas are primarily solitary and crepuscular, meaning they are most active at dawn and dusk. They are excellent climbers and spend much of their lives in trees, using their sharp claws and flexible ankles to navigate branches. In cold weather, red pandas wrap their tails around themselves like a blanket to conserve heat. 

## Conservation 

The International Union for Conservation of Nature (IUCN) lists the red panda as *Endangered*. Habitat loss, deforestation, and poaching are major threats to the species. Current estimates suggest there are fewer than 10,000 mature individuals left in the wild. Conservation efforts include habitat protection, anti-poaching initiatives, and breeding programs in zoos worldwide. 

## Cultural Significance 

Red pandas hold cultural importance in the regions where they live. In Nepal, they are called *habre*, and in Bhutan, they are considered a symbol of ecological health. Globally, red pandas have become popular ambassadors for conservation, helping raise awareness about the need to protect biodiversity in the Himalayas.

The red panda’s unique biology, charming appearance, and vulnerable status make it a species of high conservation concern. Protecting red pandas not only preserves a fascinating mammal but also helps maintain the rich forest ecosystems of the Himalayan region.

## Typical weights of a red panda 

We can use simulation to explore typical weights of red panda. @tab-red-panda-weights shows an extract of such a simulated dataset, while @fig-red-panda-weights shows a histogram of the full dataset. Here data are rounded to the nearest 0.1kg, and breakpoints have been chosen so as not to impose any smoothing. The mean weight is XX kg with a standard deviation of YY kg. 

```{r}
#| tbl-cap: Extract of simulated red panda weights.
#| label: tab-red-panda-weights
# file paths within a notebook are relative to that notebook
red_panda_weights <- read.csv("../data/red-panda-weights.csv", row.names = 1)
head(red_panda_weights)
```

```{r}
#| fig-cap: Histogram of red panda weights.
#| fig-align: center
#| label: fig-red-panda-weights
hist(
  x = red_panda_weights$weight,
  main = "",
  xlab = "weight (kg)",
  breaks = seq(3.05, 7.55, by = 0.1))
```


The mean of a sample $x = (x_1, \ldots, x_n)$ is calculated as $\bar x = \frac{1}{n}\sum_i=1^n x_i,$ while the variance is calculated as:

$$
\text{Var}(x) = \frac{1}{n-1} \sum_{i=1}^n (x_i - \bar x)^2.
$$

---
